---
title: "A Data Analysis of the New York City Airbnb"
author: " By Siddharth Sudhakar"
date: "May 8, 2020"
output:
  html_document:
    code_folding: hide
    css: style1.css
    highlight: monochrome
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float: no
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r global_options, include=FALSE, echo = False}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

## Introduction
Airbnb is an online marketplace for arranging or offering lodging, primarily homestays, or tourism experiences. It acts as a broker, receiving commissions from each booking.It currently covers more than 81,000 cities and 191 countries worldwide.<br>
The [dataset](https://www.kaggle.com/dgomonov/new-york-city-airbnb-open-data) we will be using contains 16 features about Airbnb listings within New York City.


```{r}
library(tidyverse)
library(ggridges)
library(wesanderson)
library(wordcloud)
library(RColorBrewer)
library(tm)
library(gridExtra)
data <- read_csv("AB_NYC.csv")
  theme_func <- function() {
  theme_minimal() +
  theme(
    text = element_text(family = "serif", color = "gray25"),
    plot.subtitle = element_text(size = 12,hjust = 0.5,color = "gray45"),
    plot.caption = element_text(color = "gray30"),
    plot.background = element_rect(fill = "gray95"),
    plot.margin = unit(c(5, 10, 5, 10), units = "mm"),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(color = "white")
  )
}

#na_count <-sapply(data, function(y) sum(is.na(y)))
#na_count <- data.frame(na_count)   
#na_count
names_to_delete <- c("id", "host_id","last_review")
data[names_to_delete] <- NULL
data$reviews_per_month[is.na(data$reviews_per_month)] = mean(data$reviews_per_month, na.rm=TRUE)
names_to_factor <- c("host_name", "neighbourhood_group", "neighbourhood", "room_type")
data[names_to_factor] <- map(data[names_to_factor], as.factor)



```

## Distribution of price
Initally the distribution was highly skewed to the right,by applying log transformation we get something close to a bell shaped curve.<br>
The average price among all the neighbourhood listings is found to be **152.75$**.

```{r fig.align="center"}
mean_price <- data %>%
  group_by(neighbourhood_group) %>%
  summarise(price = round(mean(price), 2))

ggplot(data, aes(price)) +

  geom_histogram( aes(y = ..density..), fill = "steelblue") + 
  geom_density(alpha = 0.1, fill = "steelblue") +
ggplot2::annotate("text", x = 1500, y = 1.0,label = "Mean price = 152.75$", size = 5) +
  geom_rug(alpha = 0.05)  +

  geom_vline(xintercept = round(mean(data$price), 2), size = 1, linetype = 2) +
  scale_x_log10() + 
  theme_func() +
        labs(
    subtitle = expression("With" ~'log'[10] ~ "transformation on x-axis"),
    title = "Distribution of price"
  )


```

## Distribution of neighbourhood groups
### By price
As we saw earlier the distribution is skewed to the right, with most of the listings have prices below 250$.<br> 
Manhattan seems to be more expensive compared to other neighbourhood groups as there are large number of listings having price more than 250$.It can also be infered that Staten Island and Bronx have relatively smaller number of listings.


```{r fig.align = "center"}


data %>% 
    mutate(neighbourhood_group = factor(neighbourhood_group, levels = c("Manhattan","Brooklyn","Queens","Bronx","Staten Island") )) %>% 
    ggplot(aes( x = price, y = neighbourhood_group)) + 
    
    geom_point(
        alpha = 0.2,
        shape = '|',
        position = position_nudge(y = -0.05)) +
    # Set bandwidth to 3.5
    geom_density_ridges(aes(fill =neighbourhood_group),bandwidth = 8.5,alpha= 0.7) +
    # add limits of 0 to 150 to x-scale
    scale_x_continuous(limits = c(0,450),expand  = c(0,0)) +
        theme( axis.ticks.y = element_blank(),
               legend.position = "none")+
    # provide subtitle with bandwidth
    labs(title = 'Distribution of neighbourhood_group by price') +
      theme_func()
```


### By price and room type

The pricing of the room type follows the order Entire home apt > Private room > Shared room,this trend is true across all the neighbourhood groups.<br>
As we found earlier Manhattan is the most expensive neighbourhood followed by Brooklyn and Queens.
```{r fig.align = "center"}
data %>%
    ggplot(aes(x = neighbourhood_group, y = price)) + 


    geom_boxplot(alpha = 0.7,aes(fill = room_type) ) +
    scale_y_log10()+
    # Reset point size to default and set point shape to 95
   geom_point(alpha = 0.2,shape = 95) +
   
     geom_hline(yintercept = mean(data$price), linetype = 2)+
    # Supply a subtitle detailing the kernel width
    labs(title = 'Distibution of neighbourhood_group by price and room_type',
             subtitle = expression("With" ~'log'[10] ~ "transformation on y-axis")) +
    theme_func()
```

### By location
The most expensive neighbourhood groups Manhattan,Queens and Brooklyn appear to be close to each other.<br>
The plot for Manhattan and Brooklyn are very dense and can be understood as being the most popular neighbourhood for Airbnb in New York.<br>
Staten Island and Bronx are not very popular for Airbnb. 

```{r fig.align = "center"}
ggplot(data, aes(longitude,latitude))+
  stat_density_2d(aes(color = neighbourhood_group)) +
  ggtitle("2-D density plot of neighbourhood_group") +
    theme_func()


```

## Top 40 Neighbourhoods
The most popular neighbourhoods are Williamsburg,Bedford-Stuyvesant,Harlem and Bushwick.
As we saw earlier the most popular neighbourhood groups are Manhattan and Brooklyn.
```{r fig.align = "center"}

data%>%
  group_by(neighbourhood,neighbourhood_group)%>%
  summarise(count = n())%>%
  arrange(desc(count))%>%
  head(30) %>%
   ggplot(aes(x = fct_reorder(neighbourhood,count),y = count,fill = neighbourhood_group ))+
  geom_col() +
  coord_flip() +

scale_fill_manual(values = wes_palette("GrandBudapest1", n = 3)) +
  theme_func() +
  ggtitle("Top 40 neighbourhoods by count") + 
  labs(x = 'neighbourhood')

```

## Relationship between price, number of reviews and minimum nights
The most expensive listings have very few or even zero reviews,and with cheaper listings having more number of reviews.<br>
The same relationship applies to minimum nights, with most expensive listing requiring lesser nights to book.

```{r fig.align = "center"}
reviews_price <- ggplot(data,aes(x = number_of_reviews,y=price))+
  geom_point(alpha = 0.2,color = "darkorchid2") +
  theme_func() +
      labs(title = 'Correlation between number_of_reviews and price')

minimum_nights_price <- ggplot(data,aes(x = minimum_nights,y=price))+
  geom_point(alpha = 0.6,color="goldenrod2") +
  theme_func() +
        labs(title = 'Correlation between minimum_nights and price')

grid.arrange(reviews_price,minimum_nights_price,nrow = 2)

```

## Wordclouds for titles
Here I wanted to check the difference in word choice of titles between Low and high price Airbnb listings.<br>
In the low price category the most popular words as titles seems to be **cozy,apartment,bedroom,village and sunny**.

```{r fig.align = "center"}
pal2 <- brewer.pal(8,"Dark2")
word_data <- data %>% 
  mutate(name = strsplit(as.character(name), " ")) %>%
  unnest(name) %>%
  group_by(name)%>%
  filter(price < 130.72)
#word_data <- slice(word_data,1:10)
word_data <- word_data[1:10000,"name"]


corp_title <- Corpus(VectorSource(word_data$name))
corp_title <- tm_map(corp_title, tolower)
title_cleaned <- tm_map(corp_title,removeWords, stopwords("english"))

tdm <- TermDocumentMatrix(title_cleaned)
m <- as.matrix(tdm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
 wordcloud(d$word,d$freq, scale=c(8,.3),min.freq=1,max.words=100, random.order=T, rot.per=.35, colors=pal2, vfont=c("sans serif","plain"))
 
```


In the expensive listings the popular words used as titles are **spacious,luxury,private,duplex and garden.**
```{r fig.align = "center"} 
 pal2 <- brewer.pal(8,"Dark2")
word_data <- data %>% 
  mutate(name = strsplit(as.character(name), " ")) %>%
  unnest(name) %>%
  group_by(name)%>%
  filter(price > 182.72)
#word_data <- slice(word_data,1:10)
word_data <- word_data[1:10000,"name"]
corp_title <- Corpus(VectorSource(word_data$name))
corp_title <- tm_map(corp_title, tolower)
title_cleaned <- tm_map(corp_title,removeWords, stopwords("english"))

tdm <- TermDocumentMatrix(title_cleaned)
m <- as.matrix(tdm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
 wordcloud(d$word,d$freq, scale=c(8,.3),min.freq=1,max.words=100, random.order=T, rot.per=.35, colors=pal2, vfont=c("sans serif","plain"))
```

## Conclusion
Manhatt

 